---
title: "Untitled"
author: "Kija"
date: "`r Sys.Date()`"
output: html_document
knitr::opts_chunk$set(echo = TRUE)
- we selected 20 isolates
- 6 isolates are ***Klebsiella spp***._ isolates
- isolates are *Escherichia coli*_ isolates 
- 8 isolates are *salmonella spp*_ isolates
here::here()
path_data <- here::here("data/2024-09-25_INIKA_SAMPLING_ANALYSIS_HUMAN.csv")
  readr::read_csv2(path_data)
editor_options: 
  chunk_output_type: console
---
# Day 3

## packages into memory
```{r}
?library()
help("library")
```

## we get out data
```{r}
path_data <- here::here("data/2024-09-25_INIKA_SAMPLING_ANALYSIS_HUMAN.csv")

path_data
```
```{r}
human_data <-
  readr::read_csv2(path_data)
```

- Names of columns

```{r}
colnames(human_data)
```

- getting the columns into a vector for pasting and manipulations

```{r}
dput(
  colnames(human_data)
  )
```

- exploration: c() is to make vectors

```{r}
test <- c("one", "2', 3")
test
test[3]
typeof(test)

class(test)
```

All the elements of a vector are of the same 'type' (text, double = Real, interger, date ...)

```{r}
test2 <-c(1,2,3)
typeof(test2)
typeof(test2 * test2)
class(test2)
```

```{r}
test2 * test2
```

```{r}
test2 * c(1,2,3)
```

- overview of data and data types
```{r}
glimpse(human_data)
str(human_data)
```



<!--
```{r}
original_names <- colnames(human_data)
original_names
```

```{r}
library(tidyverse)
new_names <- str_replace_all(original_names, " ", "_")
new_names
new_names[88]
```

```{r}
new_names2 <- str_replace_all(
  new_names, "[?/,:.*()-]", "_" ) %>%
  str_remove_all("(^_*)|(_*$)")
```

```{r}
colnames(human_data) <- new_names2
```

-->

- selecting columns to reduce the number of columns

```{r}
new_columns <- 
  colnames(human_data) %>%
  str_replace_all(" ", "_") %>%
  str_replace_all( "[?/,;.*()-]", "_") %>%
  str_remove_all("(^_*)|(_*$)")

colnames(human_data) <- new_columns
```



```{r}
human_data_selection <- 
  human_data %>%
  select(INIKA_OH_TZ_ID, Age__yrs, Gender, Enter_a_date, Region, District, 
         Specify_if_other_district, Sample, Season, Origin_of_sample, 
         Which_class_grade_are_you, 
         Who_is_your_caretaker, 
         If_others__mention, 
         What_is_your_occupation_and_or_of_your_caretaker, 
         Have_you_ever_heard_about_AMR, If_yes__how_did_you_get_this_information, 
         Have_you_or_your_children_used_any_antibiotics_at_any_time, 
         If_yes__where_did_you_get_these_drugs_from,
         If_it_was_drug_sellers_or_pharmacy__did_you_have_a_prescription_from_the_doctor_prescriber,
         GPS_coordinates_latitude, GPS_coordinates_longitude) 

human_data_selection %>% View()
```

```{r}
glimpse(human_data_selection)
str(human_data_selection)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

-> viewing your data or better an abstract of your data. Notion of pipe and selecting columns

```{r}
glimpse(human_data_selection)
str(human_data_selection)
```

# HERE
-something does not make sense - empty column


```{r}
human_data_selection <-
  human_data_selection %>%
  select(-Specify_if_other_district) %>%
  mutate_at(vars(Age__yrs), as.integer) %>%
  mutate_if(is.character, factor) %>%
  mutate_at(vars(INIKA_OH_TZ_ID), as.character)
```

- Summary

```{r}
summary(human_data_selection)
```

```{r}
human_data_selection %>%
  filter(INIKA_OH_TZ_ID == "238")
```

```{r}
human_data_selection %>%
  filter(INIKA_OH_TZ_ID != "238") %>%
  filter_(is.na(Gender))
  #summary()
```

```{r}
human_data_selection %>%
  filter(INIKA_OH_TZ_ID != "238") %>%
  filter(is.na(Gender))
#summary()
```

```{r}

```

